---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# siera

<!-- badges: start -->
<!-- badges: end -->

With siera, users ingest Analysis Results Standard - ARS (a CDISC Foundational standard) metadata and auto-generate
R scripts that, when run in with provided ADaM datasets, provide Analysis Results Datasets (ARDs).  

In order to use the readARS() function, users will need to provide the following:

1. A Functional JSON file, representing ARS Metadata for a Reporting Event (to get started, see TFL Designer)
2. An output directory where the R scripts will be placed
3. A folder containing the related ADaM datasets for the ARDs to be generated

## Installation

The current version (0.1.0) of siera can be installed from [GitHub](https://github.com/clymbclinical/siera) with:

``` {r install}
# install.packages("devtools")
Sys.unsetenv("GITHUB_PAT")
devtools::install_github("https://github.com/clymbclinical/siera")
```


## Example

Using an example JSON ARS file, run readARS() function to produce the ARD programs.

In this example, the following is used:
1. JSON ARS file: CDISC Pilot Study Common Safety Displays JSON ARS.
2. get

```{r example}
library(siera)

json_path <- system.file("extdata", "ARS_V1_Common_Safety_Displays.json", package = "siera")
output_folder <- file.path(paste0(getwd()))
ADaM_folder <- file.path(paste0(getwd(),"/ADAM"))

readARS(json_path, output_folder)
```


Once the R programs are created, they can be individually run, provided that the ADaM datasets
are in the location as provided to the readARS function. In this example, one of the 5 outputs are
Out14-1-1.  Its corresponding R script is ARD_Out14-1-1.R. 

We also assume that there are CSV ADaMs in a folder called "adam_csv" in the current working directory.

In this example, the ARD programs can be called as follows:

```{r ARD run}

# ARD_program <- file.path(paste0(getwd(),"/ARD_Out14-1-1.R"))
# source(ARD_program)
```
