---
title: "What do I do with my ARD?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. What do I do with my ARD?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(siera)
```

## Using ARDs: Options

An Analysis Results Dataset (ARD) has multiple uses.  Consisting of machine-readable results (as opposed to static RTF or PDF results), examples of ARD use cases are:

- QC datasets
- re-using results for different outputs, e.g. figures or tables
- re-using results for various reporting events
- creating final TFLs

(The [Concepts and conventions](concepts.html) vignette discusses _why_ an ARD makes these use cases easier.)

For rendering, each ARD row already carries a display-ready result: the `disp` column holds the raw statistic (`res`) formatted according to the ARS `resultPattern` for its operation (e.g. `(N=86)`, `75.2`, `( 12.5)`), so values can be placed directly into an RTF/PDF table without re-deriving decimal places or percentage scaling. See the [ARD script structure](ARD_script_structure.html) vignette for how these columns are built.

### Downstream R packages for table creation

Two reputable R packages, part of the [pharmaverse](https://pharmaverse.org/) have an ARD-first approach to creating tables:

- [`tfrmt`](https://github.com/GSK-Biostatistics/tfrmt)
- [`gtsummary`](https://www.danieldsjoberg.com/gtsummary/)

Below are presentations delivered at posit::conf 2025 on these packages:

- [tfrmt](https://posit-conf-2025.github.io/pharmaverse/slides/05-tables-tfrmt/#/title-slide)
- [gtsummary](https://posit-conf-2025.github.io/pharmaverse/slides/06-tables-gtsummary/#/title-slide)
