---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE}
library(siera)
```

We will use the following example to get started:

A modified *CDISC Common Safety Displays* ARS metadata (xlsx format, with example dynamic AnalysisMethodCodeTemplate R code handling operations, based on functions from the `cards` package)

In order to facilitate the example, _siera_ includes several example files, which we use throughout the documentation.  These include several ARS files (json and xlsx), as well as csv-format ADaMs which can be run with the R scripts produced by readARS function. Use the helper ARS_example() with no arguments to list them or call it with an example filename to get the path.

### What you will produce

Before we start, it helps to see where we are headed.  The end product of _siera_ is an **Analysis Results Dataset (ARD)** - one result per row, in machine-readable form.  Below is the ARD for a Demographics output, produced by running one of the example scripts that _siera_ generates (more on running scripts further down).  This is what we are building toward:

```{r target ARD, message=FALSE, warning=FALSE}
# Run a ready-made example script shipped with siera; the result is the object `ARD`
source(ARD_script_example("ARD_Out14-1-1.R"))
head(ARD)
```

Each row carries the statistic together with metadata (such as `AnalysisId` and the grouping columns) that traces it back to the ARS definition.  The [Concepts and conventions](concepts.html) vignette explains these traceability columns in more detail.

```{r example files}
# To see a list of example files:
ARS_example()

# A temporary path to a specific file:
ARS_example("exampleARS_1.json")
```

### Reading in ARS metadata - what to expect

The ARS metadata contains all ARS information for the Reporting Event. This information includes:

- _mainListOfContents_ (links outputs to their respective analyses)
- _otherListsOfContents_ (contains Output metadata)
- _dataSubsets_ (filters data for individual analyses)
- _analysisSets_ (filters data to get the Population Set for the Output)
- _analysisGroupings_ (groups data according to analysis specifications)
- _analyses_ (contains linking information to all components required for calculation of results)
- _methods_ (describes operations to be performed on data to get results)

For a fuller treatment of what each of these sections is for, and how they flow through _siera_, see the [Concepts and conventions](concepts.html) vignette.

### Example

The _readARS_ function reads in a completed ARS metadata file, and generates R scripts for each Output defined in the file.  Below is an example of how this can be done using a ready-to-use ARS metadata Excel file (note that this is meant as an example.  Official ARS metadata is in json format, which can also be passed to the _readARS_ function):

```{r Excel ARS metadata example, message=FALSE}
# Path to the Excel ARS metadata file:
ARS_path <- ARS_example("Common_Safety_Displays_cards.xlsx")

# Path to a folder which will contain the Output meta-programmed R scripts (recommended to update
# to a more suitable local path)
output_folder <- tempdir()

# Path to the folder containing ADaM datasets in csv format (we will use temporary
# directory tempdir() to make the code run in this vignette, but it's recommended to
# 1. download the ADaMs required (csv ADSL and ADAE available using e.g. ARS_example("ADSL.csv"))
# 2. store it in a folder (the directly downloaded location can be found using dirname(ARS_example("ADSL.csv")), for example.  Use this location, or manually store the ADaM somewhere else)
# 3. use the folder path instead of tempdir() below.
ADaM_folder <- tempdir()

# run the readARS function with these 3 parameters.  This creates R scripts
# (1 for each Output in output_folder)
readARS(ARS_path, output_folder, ADaM_folder)
```

You should now have 5 R scripts (named ARD_Out14-1-1.R, ARD_Out14-3-1-1.R, ARD_Out14-3-2-1.R, ARD_Out14-3-3-1a.R, and ARD_Out14-3-3-1b.R) in the folder specified as _output_folder_.  You can execute any of these 5 R scripts as-is (assuming the ADaM required for this script is available in the _ADaM_folder_), and the result will be an ARD (one result per row format) for each of the scripts.

The generated scripts are portable across operating systems: any Windows-style backslashes in the ADaM path are normalised to forward slashes, so a script generated on Windows runs unchanged on macOS or Linux (and vice versa).

#### Running a auto-generated ARD script
The structure of the auto-generated ARD script is discussed in the [ARD structure](ARD_script_structure.html) vignette.  Examples of such auto-generated ARD scripts are included in the package, and can be accessed with the `ARD_script_example` function. Running one (as we did to produce the preview at the top of this vignette) looks like this:

```{r run ARD_xxx, message=FALSE, warning=FALSE, eval=FALSE}
# Step 1: open ARD_xxx.R file
# Step 2: Confirm the location of ADaM dataset(s) is correct in the code section "Load ADaM".
# For the sake of simplicity, the only update made to the ARD_Out14-1-1.R script was to point to the ADaM folder to ARS_example("ADSL.csv")
# Step 3: Run the code and enjoy automated analysis results generation.
# Note: the ARD is contained in the object "ARD" (which contains the appended) mini-ARDs from all individual Analyses ARDs.

example_ARD_script <- ARD_script_example("ARD_Out14-1-1.R")
source(example_ARD_script)
head(ARD)
```

The `ARD` object is the result of appending all analysis-level ARD (created from the `cards` and `cardx` functions) for the output.  It can be used (and re-used) downstream for various applications.  See [this vignette](apply-ARD.html) for next step in utilising the ARD - like creating Tables, or using it for QC.

### A note on example code used in AnalysisMethodCodeTemplate: using `cards` package

A powerful aspect of the ARS model, is that dynamic code can be supplied as part of the metadata and be evaluated on either 1. the Output level, 2. the Analysis level, or 3. the AnalysisMethod level (for more information on this, see [this](https://www.lexjansen.com/phuse-us/2025/ds/PAP_DS09.pdf) PHUSE paper).  The latter case is implemented by making use of the _AnalysisMethodCodeTemplate_ and _AnalysisMethodCodeParameters_ classes in the metadata.  The user supplies dynamic R code (with the option of making use of parameters referring to metadata pieces), which will be executed to perform the relevant AnalysisMethod operations.  As was the case in this example, the `cards` and `cardx` R packages were used. See more information on this in the vignette on [using `cards` and `cardx`](using-cards.html).
