---
title: "Read ARS from Excel"
author: "Malan Bosman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read ARS from Excel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## ARS Metadata in Excel

The official CDISC ARS metadata is represented in JSON format, allowing language-agnostic and efficient data transfer.  However, there is an unofficial representation of ARS metadata in Excel, which has the benefit of being more easily human readable (and directly editable if needed).  An example of such an Excel ARS sheet for the CDISC Pilot study can be found [here](https://github.com/cdisc-org/analysis-results-standard/blob/main/workfiles/examples/ARS%20v1/Common%20Safety%20Displays.xlsx)

## Reading in Excel ARS metadata - what to expect

The ARS metadata represented in Excel contains all ARS information for the Reporting Event.  It is divided into Excel Worksheets, roughly mapped/equivalent to the official JSON ARS representation and its classes (an example can be found [here](https://github.com/cdisc-org/analysis-results-standard/blob/main/workfiles/examples/ARS%20v1/Common%20Safety%20Displays.json)).  For the purpose of automating Analysis Results using 
_siera_, a subset of the multiple Excel Worksheets contain the key information.  These include:
- _MainListOfContents_ (links outputs to their respective analyses)
- _OtherListsOfContents_ (contains output metadata)
- _DataSubsets_ (filters data for individual analyses)
- _AnalysisSets_ (filters data to get the Population Set for the output)
- _AnalysisGroupings_ (groups data according to analysis specifications)
- _Analyses_ (contains linking information to all components required for calculation of results)
- _AnalysisMethods_ (describes operations to be performed on data to get results)
- _AnalysisMethodCodeTemplate_ (contains dynamically written R code to perform AnalysisMethod)
- _AnalysisMethodCodeParameters_ (populates AnalysisMethodCodeTemplate with Analysis information)

### A note on example code used in AnalysisMethodCodeTemplate

A powerful aspect of the ARS model, is that dynamic code can be supplied as part of the metadata and be evaluated on either 1. the Output level, 2. the Analysis level, or 3. the AnalysisMethod level (for more information on this, see [this](https://www.lexjansen.com/phuse-us/2025/ds/PAP_DS09.pdf) PHUSE paper).  The latter case is implemented by making use of the _AnalysisMethodCodeTemplate_ and _AnalysisMethodCodeParameters_ classes in the metadata.  The user supplies dynamic R code (with the option of making use of parameters referring to metadata pieces), which will be executed to perform the relevant AnalysisMethod operations.  

For the example ARS metadata in Excel shipped with this package, the R code used in the _AnalysisMethodCodeTemplate_ is aligned with the _cards_ R package (more information [here](https://insightsengineering.github.io/cards/latest-tag/)).  The _cards_ package is a reputable R package consisting of wrappers for various statistical functions, and producing results in a consistent ARD format. Below is a simple example of how a function _ard_continuous_ from the _cards_ package is used in the _AnalysisMethodCodeTemplate_ class in the accompanying Excel ARS metadata example shipped with this package:

```{r cards example, eval=FALSE}
# example of 'cards' code in AnalysisMethodCodeTemplate, using the ard_continuous function:

Analysis_ARD = ard_continuous(
  data = filtered_ADSL,
  by = c(byvariables_here),
  variables = analysisvariable_here
)
```

In the above example of _AnalysisMethodCodeTemplate_ code for a particular AnalysisMethod (summary statistics for a continuous variable), the _ard_continuous_ function from the _cards_ package is used.  Notice, however, that two parameters are used in the code: 

1. 'byvariables_here', and 
2. 'analysisvariable_here'.  

These are example placeholder names, to be replaced by actual variables, depending on the context for the Analysis.  For example, 'byvariables_here' could be replaced by TRT01A, and 'analysisvariable_here' could be replaced by AGE, for an ARD with summary statistics on AGE by Treatment.  The same code, however, will be reused for summary statistics on e.g. HEIGHT for another Analysis, and the 'analysisvariable_here' will be replaced with HEIGHT.  All this is dynamically applied within the readARS_xl function when generating R scripts using _siera_.

## reading in Excel ARS metadata

The _readARS_xl_ function reads in a completed ARS metadata Excel file, and generates R scripts for each output defined in the file.  Below is an example of how this can be done using a ready-to-use ARS metadata Excel file:

```{r library, warning=FALSE}
library(siera)
```

```{r Excel ARS metadata example, message=FALSE}

# Path to the Excel ARS metadata file: 
ARS_path <- ARS_example("Common_Safety_Displays_cards.xlsx")

# Path to a folder which will contain the output meta-programmed R scripts (recommended to update
# to a more suitable local path)
output_folder <- tempdir()

# Path to the folder containing ADaM datasets in csv format (we will use temporary 
# directory tempdir() to make the code run in this vignette, but it's recommended to 
# 1. download the ADaMs required (csv ADSL and ADAE available using e.g. ARS_example("ADSL.csv"))
# 2. store it in a folder (the directly downloaded location can be found using dirname(ARS_example("ADSL.csv")), for example.  Use this location, or manually store the ADaM somewhere else)
# 3. use the folder path instead of tempdir() below.
ADaM_folder <- tempdir()

# run the readARS function with these 3 parameters.  This creates R scripts 
# (1 for each output in output_folder)
readARS_xl(ARS_path, output_folder, ADaM_folder)
```

### Try it out!

If you've followed along in the previous section example, you should now have 5 R scripts (named ARD_Out14-1-1.R, ARD_Out14-3-1-1.R, ARD_Out14-3-2-1.R, ARD_Out14-3-3-1a.R, and ARD_Out14-3-3-1b.R) in the folder specified as _output_folder_.  If this is successful, you can execute any of these 5 R scripts as-is (assuming the ADaM required for this script is available in the _ADaM_folder_), and the result will be an ARD (one result per row format) for each of the scripts.  This will look like this:



Note: each output requires one or more ADaM datasets to calculate the results from.  The code to load these csv ADaMs are automatically added to the ARD_xxx.R script.  As mentioned, some ADaMs are available using the ARS_example("XXXX.csv") function. However, Output 14-3-3-1a and Output 14-3-3-1b make use of the ADVS dataset, which is not currently shipped with the _siera_ package due to size considerations.  The relevant ADVS ADaM van can be obtained from the CDISC Pilot Study.  All other outputs use either ADSL or ADAE, which is shipped with this package and can be accessed through
